---
title: "Extract  data"
author: "Dave Fifield"
date: "`r format(Sys.time(), '%d %B, %Y %H:%M:%S')`"
output: 
  html_document: 
    toc: yes
    number_sections: yes
editor_options: 
  chunk_output_type: console
---

```{r setup}
source(here::here("R/analysis_settings.R"), echo = T)
```

# Import Ship-based data

## Extract ECSAS Survey data

Extract all data for ECSAS ship-based surveys. This will get everyting including
data in Pacific and Eastern Atlantic. I can do analysis for entire dataset and
then subset to areas of interest later.


```{r extract_ECSAS_ship}
if (ECSAS.reextract) {
  ecsas.ship.dat <- ECSAS.extract(
    distMeth = "All",
    intransect.only = FALSE, # Lets get everything in ECSAS
    years = c(2006, 2024),
    ecsas.path = ECSAS.Path
  )  

  write_rds(ecsas.ship.dat,
            file = file.path(GenDataDir, "ECSAS_ship_raw_data.rds"))
} else {
  ecsas.ship.dat <- read_rds(file.path(GenDataDir, "ECSAS_ship_raw_data.RDS")) 
}

progs <- ECSAS.get.table(ecsas.path = ECSAS.Path, "lkpProgram")
ecsas.ship.dat <- ecsas.ship.dat %>%
  left_join(progs, by = c("Program" = "ProgramID")) %>%
  dplyr::filter(
    Program %in% c(4, 6),
    DataSharing == 1
  ) %>%
  mutate(
    # SurveyType = "Ship",
    # spec.grp = find.spec.grp(Alpha),
    # InTransect = case_when(InTransect == -1 ~ TRUE, 
    #                        InTransect == 0 ~ FALSE, 
    #                        .default = NA),
    # Distance = Distance / 1000,
    # Windspeed = case_when(
    #   is.na(Windspeed) &
    #     !is.na(Windforce) ~ left_join(., beaufort_conversion, 
    #                                   by = c("Windforce" = "beaufort"))$speed.kts,
    #   TRUE ~ Windspeed
    # ),
    # Visibility = case_when(Visibility > 20 ~ 20, TRUE ~ Visibility),
    # FlySwim = as.factor(toupper(FlySwim)),
    # distbegin = NA_integer_,
    # distend = NA_integer_,
    # BlindWidth = NA,
    WatchID = paste0("ECSAS_shp_", WatchID),
  )

ecsas.ship.watches <- ecsas.ship.dat %>% 
  select(WatchID, LatStart, LongStart, Date) %>% 
  distinct()

ecsas.ship.distdata <- ecsas.ship.dat %>%
  dplyr::filter(
    !is.na(Count),
    !is.na(DistMeth),
    !is.na(FlockID),
    FlySwim %in% c("W", "F"),
    DataSharing == 1
  ) %>%
  mutate(
    SurveyType = "Ship",
    spec.grp = find.spec.grp(Alpha),
    InTransect = case_when(InTransect == -1 ~ TRUE, 
                           InTransect == 0 ~ FALSE, 
                           .default = NA),
    Distance = Distance / 1000,
    # Distance = case_when(is.na(Distance) ~ 0, TRUE ~ Distance),
    DistType = assign.dist.type(.),
    Windspeed = case_when(
      is.na(Windspeed) &
        !is.na(Windforce) ~ left_join(., beaufort_conversion, 
                                      by = c("Windforce" = "beaufort"))$speed.kts,
      TRUE ~ Windspeed
    ),
    Visibility = case_when(Visibility > 20 ~ 20, TRUE ~ Visibility),
    FlySwim = as.factor(FlySwim),
    distbegin = NA_integer_,
    distend = NA_integer_,
    BlindWidth = NA,
  ) %>%
  rename(object = FlockID,
         size = Count,
         distance = Distance) %>%
  select(
    WatchID,
    object,
    Alpha,
    Association,
    BlindWidth,
    DataSharing,
    Date,
    distance,
    distbegin,
    distend,
    DistMeth,
    DistType,
    English,
    FlySwim,
    InTransect,
    LatStart,
    LongStart,
    Seabird,
    SeaState,
    size,
    spec.grp,
    SurveyType,
    Swell,
    Waterbird,
    Windspeed,
    Year
  ) %>%
  droplevels


# Summary
summary(ecsas.ship.watches)
summary(ecsas.ship.distdata)

ecsas.ship.the.data <- list(distdata = ecsas.ship.distdata, 
                            watches = ecsas.ship.watches)
write_rds(ecsas.ship.the.data,
     file = file.path(GenDataDir, "ECSAS_ship_the_data.rds"))
# ecsas.ship.the.data <- read_rds(file = file.path(GenDataDir,"ECSAS_ship_the_data.rds"))
```

# Import Aerial Data

## Get ECSAS aerial watches

``` {r ECSAS_aerial_watches}
# This was created by copying ECSAS_aerial_watches.rda from Atlantic DSM
# analysis and re-saving as an rds file.
ecsas.air.watches <- read_rds(file = file.path(GenDataDir, "ECSAS_aerial_watches.rds"))
```

## Process ECSAS aerial on-effort data

```{r ecsas_air_on_effort}
# Import aerial data remove QC data for now since we have no positions for them.
ecsas.air.dat.on.effort <- readxl::read_xlsx(file.path(RawDataDir, "q Aerial on-effort data.xlsx")) %>%
  filter(
    !is.na(AerialObsID),
    StudyAreaName != "GenericPQ",
    FlySwim %in% c("W", "F"),
    !is.na(AerialObsID),
    DataSharing == 1
  )  %>%
  select(
    AerialObsID,
    AerialTransectID,
    StartDateTime,
    EndDateTime,
    Alpha,
    Association,
    BlindWidth,
    Count,
    DataSharing,
    Distance_m,
    dist_begin,
    dist_end,
    DistMeth,
    English,
    FlySwim,
    InTransect,
    ObsLat,
    ObsLong,
    ObsTime,
    SeaState,
    Seabird,
    Waterbird
  ) %>%
  left_join(
    select(
      ecsas.air.watches,
      WatchID,
      StartTime,
      EndTime,
      LatStart,
      LongStart
    ),
    join_by(between(ObsTime, StartTime, EndTime, bounds = "[)")),
    relationship = "many-to-one"
  ) %>%
  mutate(Date = StartTime) %>%
  select(-StartTime,
         -EndTime,
         -EndDateTime,
         -StartDateTime,
         -ObsLong,
         -ObsLat)

# 7128 records
# 

# The foregoing may leave obs that came at the last second of the last watch before an
# off-effort break or end of transect with no matching transect because of the
# bounds = "[)" arg. So match these up using bounds = "[]". Any still without a
# matching watchid must have occurred during a pause/resume (since dbase doesn't
# check for this yet) or on problem watches, or outside of transect (which
# shouldn't happen), so filter these out.
if (any(is.na(ecsas.air.dat.on.effort$WatchID))) {
  unmatched <- filter(ecsas.air.dat.on.effort, is.na(WatchID)) %>% 
    select(-WatchID, -LatStart, -LongStart) %>% 
    left_join(select(ecsas.air.watches, WatchID, StartTime, EndTime,
                     LatStart, LongStart),
            join_by(between(ObsTime, StartTime, EndTime, bounds = "[]")),
            relationship = "many-to-one") %>% 
    filter(!is.na(WatchID)) %>% # drop off-effort or ones outside study area
    select(-StartTime, -EndTime)
  matched <- filter(ecsas.air.dat.on.effort, !is.na(WatchID))
  ecsas.air.dat.on.effort <- rbind(matched, unmatched) %>% 
    arrange(WatchID, ObsTime)
}

# No longer needed
ecsas.air.dat.on.effort <- select(ecsas.air.dat.on.effort, -ObsTime)

# 6579 rows 
summary(ecsas.air.dat.on.effort)
dftools:::namesort(ecsas.air.dat.on.effort)
```

## Process ECSAS aerial off-effort data

```{r ecsas_air_off_effort}
ecsas.air.dat.off.effort <- readxl::read_xlsx(file.path(RawDataDir, "q Aerial off-effort data.xlsx")) %>%
  filter(StudyAreaName != "GenericPQ", # FlySwim %in% c("W", "F"),
         DataSharing == 1) %>% 
  select(
    AerialObsID,
    Alpha,
    Association,
    Count,
    DataSharing,
    Distance_m,
    dist_begin,
    dist_end,
    DistMeth,
    English,
    FlySwim,
    InTransect,
    ObsLat,
    ObsLong,
    ObsTime,
    Seabird,
    Waterbird
  ) %>% 
  mutate(
    SeaState = NA,
    BlindWidth = NA,
    AerialTransectID = NA,
    # Need a watchID to match other data, just make one up since off_effort
    # obs don't have one
    WatchID = paste(
      "ECSAS_air_off_effort",
      1:nrow(.),
      sep = "_"
    )
  ) %>% 
  rename(LatStart = ObsLat,
         LongStart = ObsLong,
         Date = ObsTime)

# 153 obs

dftools:::namesort(ecsas.air.dat.on.effort)
dftools:::namesort(ecsas.air.dat.off.effort)
```

## Combine aerial on- and off-effort data

```{r combine_ecsas_aerial}
ecsas.air.distdata <- rbind(ecsas.air.dat.on.effort, ecsas.air.dat.off.effort) %>%
  rename(
    distbegin = dist_begin,
    distend = dist_end,
    object = AerialObsID, #TODO:Possible to clash with on_effort AerialobsID??
    size = Count,
    distance = Distance_m
  ) %>%
  filter(!is.na(LatStart), !is.na(LongStart)) %>%  # Some off-effort have none
  mutate(
    SurveyType = "Air",
    DistType = "Perp.",
    Swell = NA,
    Windspeed = NA,
    distbegin = distbegin / 1000,
    distend = distend / 1000,
    spec.grp = find.spec.grp(Alpha),
    Year = year(Date),
    # TODO: why?? Should leave as NA??
    # distance = case_when(is.na(distance) ~ 0, TRUE ~ distance)
  ) %>% 
  select(-AerialTransectID)

summary(ecsas.air.watches)
summary(ecsas.air.distdata)

ecsas.air.the.data <- list(
  distdata = ecsas.air.distdata,
  watches = ecsas.air.watches %>% select(WatchID, LatStart, LongStart, Date)
)

write_rds(ecsas.air.the.data,
     file = file.path(GenDataDir, "ECSAS_air_the_data.rds"))
# ecsas.air.the.data <- read_rds(file = file.path(GenDataDir,"ECSAS_air_the_data.rds"))
```

## Create final ECSAS data

```{r ecsas_combine_all_data}
# Create combined ship/air study.area

the.data <- list(
  distdata = rbind(ecsas.ship.the.data$distdata, ecsas.air.the.data$distdata),
  watches = rbind(ecsas.ship.the.data$watches, ecsas.air.the.data$watches)
)

write_rds(the.data, file = the.data.file)
# load(the.data.file)
```

## Save Shapefiles

```{r save_shapefile}

# Save shapefile
the.data$distdata %>%
  mutate(object = as.character(object)) %>%  # shapefile can't handle big ints
  st_as_sf(
    coords = c("LongStart", "LatStart"),
    crs = 4326,
    remove = FALSE
  ) %>%
  st_write(
    dsn = ShapeDir,
    layer = "ECSAS_distdata.shp",
    driver = "ESRI Shapefile",
    delete_layer = TRUE
  )

watches.sf <- the.data$watches %>%
  mutate(object = as.character(WatchID)) %>%  # shapefile can't handle big ints
  st_as_sf(
    coords = c("LongStart", "LatStart"),
    crs = 4326,
    remove = FALSE
  ) %>%
  st_write(
    dsn = ShapeDir,
    layer = "ECSAS_watches.shp",
    driver = "ESRI Shapefile",
    delete_layer = TRUE
  )
```


# Create study area and save

Study area is just MCP of all points.

```{r ECSAS_study_area}
study.area <- watches.sf %>%
  mutate(ID = 1) %>%
  select(ID, geometry) %>%
  as_Spatial() %>%
  adehabitatHR::mcp(percent = 100) %>%
  as("sf") %>%
  st_write(
    dsn = ShapeDir,
    layer = "study_area.shp",
    driver = "ESRI Shapefile",
    delete_layer = TRUE
  )

write_rds(study.area, study.area.file)
```


# Plot watches

```{r plot_watches}
dftools::plot_sf_obj(watches.sf, "LongStart", "LatStart")
```


# SessionInfo
```{r}
sessioninfo::session_info()
```
