---
title: "Create aggregated ECSAS abundance points"
author: "Dave Fifield"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---


Sum detection-corrected abundance by survey segment and convert to density. Then
rasterize to the generic grid (summing all survey segments that occur in each
grid cell). Convert the resulting raster to points. This only allows one value
per grid cell. If I want multiple attributes I'll probably need to convert the
generic grid to polygon so each one can have multiple attributes.


```{r setup}
source(here::here("R/analysis_settings.R"), echo = T)
the.data <- readRDS(the.data.file)
study.area <- readRDS(study.area.file)
proj <- readRDS(proj.file)
grid <- readRDS(grid.file)
```


# Get data

```{r get_data}
studyarea.proj <- study.area %>% st_transform(proj)
watches <- the.data$watches %>% 
  assign.season(seasons, datefield = "Date")
```


# Aggregate 

Aggregate observations by watch and then to gridcell on a seasonal basis
combining across all species of interest.

Includes two pseudo-species, SBRD and WBRD, which include all species marked
as "Seabirds" and "Waterbirds" in ECSAS, respectively.

```{r aggregate_to_raster}
species <- c("Sbrd", "Wbrd", names(spec.grps))
agg.data <- species %>%
  # Aggregate to raster cells
  map(\(grp) agg_species_group(grp, the.data$distdata, watches, grid)) %>%
  setNames(species) %>%
  # Combine list of species (each with seasonal data) to single structure
  # for each season across all species
  combine_species() %>%
  save_seasonal_shapefiles()

saveRDS(agg.data, file.path(GenDataDir, "agg_data.rds"))
# agg.data <- readRDS(file.path(GenDataDir, "agg.data.by.species.rds"))
```
