---
title: "Create aggregated ECSAS abundance points"
author: "Dave Fifield"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---


Aggregate observations to grid

Rasterize all observations to the generic grid (summing all obs that
occur in each grid cell). Convert the resulting raster to points. This only
allows one value per grid cell. If I want multiple attributes I'll probably need
to convert the generic grid to polygon so each one can have multiple attributes.


```{r setup}
source(here::here("R/analysis_settings.R"), echo = T)
```


# Get data and assign seasons

Should assign seasons in Extract_data.Rmd

```{r get_data}
the.data <- readRDS(the.data.file)
proj <- readRDS(proj.file)
grid <- rast(file.path(RasterDir, "grid.tif"))
studyarea.proj <- readRDS(study.area.file) %>% st_transform(proj)
watches <- the.data$watches %>% 
  assign.season(seasons, datefield = "Date")
```


# Aggregate 

Aggregate observations to grid cell on a seasonal basis combining across all
species of interest.

Includes two pseudo-species, SBRD and WBRD, which include all species marked
as "Seabirds" and "Waterbirds" in ECSAS, respectively.

```{r aggregate_to_raster}
species <- c("Sbrd", "Wbrd", names(spec.grps))
agg.data <- species %>%
  # Aggregate to raster cells
  map(\(grp) agg_species_group(grp, the.data$distdata, watches, grid)) %>%
  setNames(species) %>%
  # Combine list of species (each with seasonal data) to single structure
  # for each season across all species
  combine_species() %>%
  save_seasonal_shapefiles()

saveRDS(agg.data, file.path(GenDataDir, "agg_data.rds"))
# agg.data <- readRDS(file.path(GenDataDir, "agg.data.by.species.rds")) 
```
