---
title: "ECSAS observation summary"
author: "Dave Fifield"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---


```{r get_data, include=FALSE}
source(here::here("R/analysis_settings.R"), echo = FALSE)
the.data <- readRDS(the.data.file)
```

# Summary by groups

## List of multi-species groups

The following multi-species groups are used in the summary table below.

```{r single_species_groups, echo=FALSE}
library(htmltools)

# Do species groups of interest - get multi-species groups
idx <- lapply(spec.grps, length) %>% unlist() != 1
generate_html_species_list(spec.grps[idx])
```


```{r grp, echo=FALSE}

```


## Group summary table

Summary for species groups of interest.

```{r summ_grps, echo = FALSE}
# Deal with Sbrds and Wbrds first
sbrds.summ <- the.data$distdata %>%
  rename(Count = size) %>% 
  filter(Seabird == -1) %>% 
  summarise(`4_letter_code` = "Sbrd",
            Seabird = TRUE,
            Waterbird = NA,
            Num_Obs = n(),
            Num_watch = length(unique(WatchID)),
            Total_Count = sum(Count, na.rm = TRUE)) %>%
  mutate(English = "All Seabirds") %>% 
  relocate(English)

wbrds.summ <- the.data$distdata %>%
  rename(Count = size) %>% 
  filter(Waterbird == -1) %>% 
  summarise(`4_letter_code` = "Wbrd",
            Seabird = NA,
            Waterbird = TRUE,
            Num_Obs = n(),
            Num_watch = length(unique(WatchID)),
            Total_Count = sum(Count, na.rm = TRUE)) %>%
  mutate(English = "All Waterbirds") %>% 
  relocate(English)

# Now do the multi-species spec.grps
othr.grps.summ <- the.data$distdata %>%
  rename(Count = size) %>% 
  filter(!is.na(spec.grp) & spec.grp %in% names(spec.grps[idx])) %>% 
  group_by(spec.grp) %>% 
  summarise(`4_letter_code` = unique(spec.grp),
            Seabird = if (unique(Seabird) == -1) TRUE else FALSE,
            Waterbird = if (unique(Waterbird) == -1) TRUE else FALSE,
            Num_Obs = n(),
            Num_watch = length(unique(WatchID)),
            Total_Count = sum(Count, na.rm = TRUE)) %>% 
  mutate(English = unlist(spec.grps.names[spec.grp])) %>% 
  select(-spec.grp) %>% 
  relocate(English)

rbind(sbrds.summ, wbrds.summ, othr.grps.summ) %>%
  arrange(desc(Total_Count)) %>%
  knitr::kable("html") %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = TRUE
  )

```

# Individual species (or taxa) summary

These taxa are generally single species except for cases where identification
to species was not possible, in which case they are to the lowest possible
taxon. 

```{r summary_all, echo=FALSE}
# Now summarize by "English" species name from ECSAS, add in the group
# summaries from above

the.data$distdata %>%
  rename(Count = size) %>% 
  group_by(English) %>%
  summarise(`4_letter_code` = unique(Alpha),
            Seabird = if (unique(Seabird) == -1) TRUE else FALSE,
            Waterbird = if (unique(Waterbird) == -1) TRUE else FALSE,
            Num_Obs = n(),
            Num_watch = length(unique(WatchID)),
            Total_Count = sum(Count, na.rm = TRUE)) %>%
  arrange(desc(Total_Count)) %>% 
  knitr::kable("html") %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = TRUE
  )
```
